Mechanical Detection of Dipole-Dipole Interactions \\between Electronic Spins in a Solid

### 2020-10-07 Mes papiers/Méca CR/Meca_cross.tex
tutu

This effect has been routinely observed in bulk materials, the origin of which was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}. The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$. When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx ms$), the latter can drop 
to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
This process has recently revived interest recently because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
%In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
%To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.
In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias}$ of a few 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet. 
Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
For these measurement we use GREEN POWER ? REPOLARISATION RATE ? 
Three dips in the fluorescence can be observed. 
We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra at the values ...
There is indeed a strong correspondance between the PL dips and the degeneracy conditions. 
(See SI for details on planes that cross etc.). 
We tentatively attrtibute those to CR. 
%It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
%where, to our knowledge, such cross-relaxations have not been observed. 
%Indeed, surface effects may start to play a role and change quantitatively. 
To confirm our interpretation, and to extract the spin-torque dynamics in the mechanical experiment, we measure the longitudinal relaxation time ($T_1$ time) of the NV spins under different degeneracy conditions.

The measurement consists in applying a green laser to polarise the NV centers and to measure the photoluminescence at a later time. 
Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use a sequence presented in the SI, where a microwave pulse is applied or not prior to relaxation. Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
The results of the measurements are shown in Fig. \label{CR_deposited}b) for different degeneracies. 
It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.

\begin{figure}[!ht]
  \centering \scalebox{0.13}{\includegraphics{dilutednv.eps}}
  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond cristalline structure. Here, the magnetic field is tuned so that no cross-relaxation can occur.
The resulting non-resonant spin-torque $\tau_{NR}$ applied to the levitating diamond is sketched below. 
ii) Here, the magnetic field is tuned so that cross-relaxation can occur between two classes of NVs.
The resulting resonant spin-torque $\tau_{R}$ applied to the levitating diamond (below) is thus modified. 
  }\label{principle}
\end{figure}

>>>>

This effect has been routinely observed in bulk materials, the origin of which was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}. The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$. When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx ms$), the latter can drop 
to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
This process has recently revived interest recently because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
%In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
%To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.
In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias}$ of a few 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet. 
Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
For these measurement we use GREEN POWER ? REPOLARISATION RATE ? 
Three dips in the fluorescence can be observed. 
We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra at the values ...
There is indeed a strong correspondence between the PL dips and the degeneracy conditions. 
(See SI for details on planes that cross etc.). 
We tentatively attribute those to CR. 
%It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
%where, to our knowledge, such cross-relaxations have not been observed. 
%Indeed, surface effects may start to play a role and change quantitatively. 
To confirm our interpretation, and to extract the spin-torque dynamics in the mechanical experiment, we measure the longitudinal relaxation time ($T_1$ time) of the NV spins under different degeneracy conditions.

The measurement consists in applying a green laser to polarize the NV centers and to measure the photoluminescence at a later time. 
Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use a sequence presented in the SI, where a microwave pulse is applied or not prior to relaxation. Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
The results of the measurements are shown in Fig. \label{CR_deposited}b) for different degeneracies. 
It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.

\begin{figure}[!ht]
  \centering \scalebox{0.13}{\includegraphics{dilutednv.eps}}
  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond crystalline structure. Here, the magnetic field is tuned so that no cross-relaxation can occur.
The resulting non-resonant spin-torque $\tau_{NR}$ applied to the levitating diamond is sketched below. 
ii) Here, the magnetic field is tuned so that cross-relaxation can occur between two classes of NVs.
The resulting resonant spin-torque $\tau_{R}$ applied to the levitating diamond (below) is thus modified. 
  }\label{principle}
\end{figure}
